# Machine Learning Class Project Writeup

This analysis produces a model to predict the manner of performance of a fitness exercise based on sensor data.  The manner of exercise is labeled as A,B,C,D or E, with A being "correct."  See http://groupware.les.inf.puc-rio.br/har for more infomration about the study and the data set.

## Download data sets
```{r, cache=T}
trainfile <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testfile <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

download.file(trainfile,destfile="data/train.csv",method="curl")
download.file(testfile, destfile="data/test.csv",method="curl")
```

## Load data 
```{r}
trntst <- read.csv("data/train.csv")
# we'll call the "test" data "submission data"
# as we have no labels on those data and we
# won't be able to "test" or "validate" or model
# on them
sub <- read.csv("data/test.csv")
dim(trntst)
dim(sub)
```

## Create another test partition
Create a partition of labeled data to test the models on.
```{r}
library(caret)
set.seed(3000)
intrn <- createDataPartition(y=trntst$classe,p=0.7,list=F)
trn <- trntst[intrn,]
tst <- trntst[-intrn,]
```

## Try single variable classification on each variable
To start, plot predictors against "classe" to see if any 
variables separate nicely by classe.  

## Some example plots:
```{r}
set.seed(4545)
#sample the data for plotting
trn.small <- trn[sample(nrow(trn),250),]
par(mfrow=c(2,2))
plot(trn.small$classe,trn.small$X)
plot(trn.small$classe,trn.small[,41])
plot(trn.small$classe,trn.small[,83])
plot(trn.small$classe,trn.small[,93])
```

Turns out that "X" looks good (column 1). 
Running out of time as experiements take a long time.  
Will build a model using X as it has the best separation.

Note: I rejected timing variables and usernames as those don't seem to be helpful in general, although they seem to be valuable predictors in the labeled data. In fact **raw_timestamp_part_1** turned out to predict the correct outcome with 99% accuracy.  My guess is that's due to the time ordering of the experiment. 

```{r,cache=T}
fit1 <- train(classe ~ X, data=trn,method="rpart")
accuracy <- fit1$results[1,2]
accuracy
```

## Test the model on the test partition
```{r}
sum((predict(fit1,newdata=tst) == tst$classe) * 1) / length(tst$classe)
```

The accuracy drops on the test partition, but it still does fairly well.  Since I'm running out of time, I will settle with this model.  


## Predictions
```{r}
preds <- predict(fit1,newdata=sub)
preds
```

